---
title: "Covid91 vaccine study Final2021"
author: "Leslie Johnson"
date: "Summer 2021"
output:
   html_document:
         toc: true
         toc_depth: 5
---


```{r include=FALSE,echo=FALSE}
require(tidyverse)
require(tigerstats)
require(forcats)
OldData <-read.csv(file="covid91.csv",header=TRUE)
#recode to remove the annoying "X" from the dataframe
FinalData <- OldData %>% mutate(infected = recode(infected, aids = "covid", none = "OK" )) %>% select(-one_of("X"))
FinalData <- FinalData %>%
mutate(across(where(is.character),as.factor))
FinalData$treatment <-relevel(FinalData$treatment,ref = "placebo")
#FinalData$infected <- relevel(FinalData$infected,ref="OK")

```

```{r}
guy = filter(FinalData, sex =="guy")
gal = filter(FinalData, sex == "gal")
gay =filter(FinalData, LGBTQ == "gay")
Druguser = filter(FinalData,DrugUser =="yes")

```


# Introduction

A new  preventative vaccine (covid91) is under test in clinical trials. Given the nature of this vaccine it is suspected that the drug may exhibit different effectiveness across the variety of subpopulations tracked. Consider population subgroups by gender( guys and gals), drug use, and sexual orientation. Write a complete report that separately considers this drug for each subgroup and your recommendations as to whether people in the subgroup should take the vaccine. Structure your report in four sections each devoted to one of the four subgroups, use null and alternative hypothesis statements and correct statistical verbage when referencing p-values in conclusions.

### Coding Details

The variables from FinalData needs to be changed from the character variables into factor variables. Subgroups had to be separated into categories, Guys, Gals, Gays, and Drugusers to see which groups would take the COVID-19 vaccine.

# Guys

### barplot

```{r}
dd2<- guy %>% group_by(infected,treatment) %>% summarize(count=n())%>% 
mutate(prcnt=count/sum(count))

# the group_by followed by summarize(count=n())

basicC <- ggplot(dd2,aes(x=infected,y=count,fill=treatment))
basicC + geom_bar(stat="identity", position = "dodge")

# Now for percentage  plot

basicC <- ggplot(dd2,aes(x=infected,y=prcnt*100,fill=treatment))
basicC + geom_bar(stat = "identity",position = "dodge")
```

This barplot shows that around 57% took the vaccine and still contracted COVID-19, while 43% took the placebo and contracted COVID-19. The vaccine seems to be ineffective.

### Numerical results

```{r}
table1 <- xtabs(~infected + treatment, data=guy)
table2 <- xtabs(~infected + treatment, data = guy)
rowPerc(table1)
colPerc(table1)
```


### Inferential results

```{r}
chisq.test(table1)
chisqtestGC(table1)
guysfish <- fisher.test(table1)
guysfish
```

The Fisher's Exact test says the p_value above is too small at 6e-04, so we can be rejected in favor of the alternative. 

The null hypothesis predicted that the deaths related to COVID-19 are independent of sex, and it is evident that sex does not play a role in the deaths of people.

```{r}
effect <- guysfish$p.value/(guysfish$p.value +1)
effectivenessguy = guysfish$estimate/(guysfish$estimate+1)*100
print(effectivenessguy)
```

This shows us the effect odds ratio is 40.4

```{r}
low=(guysfish$conf.int[1]/(guysfish$conf.int[1] +1))*100
hi=(guysfish$conf.int[2]/(guysfish$conf.int[2] +1))*100
options
(digits = 2)
```

```{r}
guysfish <- fisher.test(table1)
guysfish
```

## Effectiveness Confidence Interval `r low` to `r hi`


# Gal

### barplot

```{r}
dd2<- gal %>% group_by(infected,treatment) %>% summarize(count=n())%>% 
mutate(prcnt=count/sum(count))

# the group_by followed by summerize(count=n())

basicC <- ggplot(dd2,aes(x=infected,y=count,fill=treatment))
basicC + geom_bar(stat="identity", position = "dodge")

# Now for percentage  plot

basicC <- ggplot(dd2,aes(x=infected,y=prcnt*100,fill=treatment))
basicC + geom_bar(stat = "identity",position = "dodge")


```

This plot shows us the population of gals by taking the COVID-19 vaccine compared to those who will take the placebo. 

### Numerical Results

```{r}
table1 <- xtabs(~infected + treatment, data=gal)
table2 <- xtabs(~infected + treatment, data=gal)
rowPerc(table1)
colPerc(table1)
```

This shows specific percentages of gals that contracted COVID-19 after taking the placebo, and the vaccine. 33% of the gals who took the vaccine still got COVID-19 compared to the 66% of gals who took the placebo and still contracted COVID-19.  

## Inferential  

```{r}
chisq.test(table1)
chisqtestGC(table1)
galsfish <- fisher.test(table1)
galsfish
```

The p_value is so low around 0.002 so the results are not likely it to occur  so the null hypothesis can be rejected.  

```{r}
effect <- galsfish$p.value/(galsfish$p.value +1)
effectivenessgal = galsfish$estimate/(galsfish$estimate+1)*100
print(effectivenessgal)
```

The effects odds shown above gives us approximately around 67%.  

```{r}
low=(galsfish$conf.int[1]/(galsfish$conf.int[1] +1))*100
hi=(galsfish$conf.int[2]/(galsfish$conf.int[2] +1))*100
options
(digits = 2)
```

```{r}
galsfish <- fisher.test(table1)
galsfish
```

## Effectiveness Confidence Interval `r low` to `r hi`



# Gay Sexual Oreintation

### barplot

```{r}
dd2<- gay %>% group_by(infected,treatment) %>% summarize(count=n())%>% 
mutate(prcnt=count/sum(count))

# the group_by followed by summarize(count=n())

basicC <- ggplot(dd2,aes(x=infected,y=count,fill=treatment))
basicC + geom_bar(stat="identity", position = "dodge")

# Now for percentage  plot

basicC <- ggplot(dd2,aes(x=infected,y=prcnt*100,fill=treatment))
basicC + geom_bar(stat = "identity",position = "dodge")

```

This barplot shows the population of gays of those who contracted COVID-19 by either taking the vaccine oppose to those who took the placebo.

### Numerical Results

```{r}
table1 <- xtabs(~infected + treatment, data=gay)
table2 <- xtabs(~infected + treatment, data=gay)
rowPerc(table1)
colPerc(table1)
```

The gays that took the vaccine and contracted COVID-19 was around 71% compared to the 29% who only took the placebo and contracted COVID-19. So the ones who took the vaccine were 3 times likely to get COVID-19 than ones that took the placebo.

### Inferential Results

```{r}
chisq.test(table1)
chisqtestGC(table1)
gaysfish <- fisher.test(table1)
gaysfish
```




```{r}
effect <- gaysfish$p.value/(gaysfish$p.value +1)
effectivenessgay = gaysfish$estimate/(gaysfish$estimate+1)*100
print(effectivenessgay)
```

```{r}
low=(gaysfish$conf.int[1]/(gaysfish$conf.int[1] +1))*100
hi=(gaysfish$conf.int[2]/(gaysfish$conf.int[2] +1))*100
options
(digits = 2)
```

```{r}
gaysfish <- fisher.test(table1)
gaysfish
```


## Effectiveness Confidence Interval `r low` to `r hi`


# Drug Users

### barplot

```{r}
dd2<- Druguser %>% group_by(infected,treatment) %>% summarize(count=n())%>% 
mutate(prcnt=count/sum(count))

# the group_by followed by summerize(count=n())

basicC <- ggplot(dd2,aes(x=infected,y=count,fill=treatment))
basicC + geom_bar(stat="identity", position = "dodge")

# Now for percentage  plot

basicC <- ggplot(dd2,aes(x=infected,y=prcnt*100,fill=treatment))
basicC + geom_bar(stat = "identity",position = "dodge")


```

This barplot represents the population of Drugusers from either taking the vaccine or the placebo. 

### Numerical Results

```{r}
table1 <- xtabs(~infected + treatment, data=Druguser)
table2 <- xtabs(~infected + treatment, data=Druguser)
rowPerc(table1)
colPerc(table1)
```

This shows that 34% who took the vaccine contracted COVID-19 compared to 66% who took the placebo and contracted COVID-19. Which makes the Drugusers half as likely to get COVID-19.     


### Inferential Results

```{r}
chisq.test(table1)
chisqtestGC(table1)
Drugusersfish <- fisher.test(table1)
Drugusersfish
```

The p_value is a small number at 0.002 so it won't likely occur by chance. The drug is effective so the null hypothesis is rejected.

```{r}
effect <- Drugusersfish$p.value/(Drugusersfish$p.value +1)
effectivenessdrug = Drugusersfish$estimate/(Drugusersfish$estimate+1)*100
print(effectivenessdrug)
```

```{r}
low=(Drugusersfish$conf.int[1]/(Drugusersfish$conf.int[1] +1))*100
hi=(Drugusersfish$conf.int[2]/(Drugusersfish$conf.int[2] +1))*100
options
(digits = 2)
```

```{r}
Drugusersfish <- fisher.test(table1)
Drugusersfish
```

## Effectiveness Confidence Interval `r low` to `r hi`



# Discussion and Conclusion 

The COVID-19 vaccine was not effective in protecting guys against contracting the disease opposed to protecting gals. COVID-19 was successful in protecting gals. If you are a guy or identify as gay it is not recommended to take the vaccine, but you are a gal or druguser the vaccine was reported to be effective.
