---
title: "Covid91 vaccine study Final2021"
author: "Leslie Johnson"
date: "Summer 2021"
output:
   html_document:
         toc: true
         toc_depth: 5
---


```{r include=FALSE,echo=FALSE}
require(tidyverse)
require(tigerstats)
require(forcats)
OldData <-read.csv(file="covid91.csv",header=TRUE)
#recode to remove the annoying "X" from the dataframe
FinalData <- OldData %>% mutate(infected = recode(infected, aids = "covid", none = "OK" )) %>% select(-one_of("X"))
FinalData <- FinalData %>%
mutate(across(where(is.character),as.factor))
FinalData$treatment <-relevel(FinalData$treatment,ref = "placebo")
#FinalData$infected <- relevel(FinalData$infected,ref="OK")

```

```{r}
guy = filter(FinalData, sex =="guy")
gal = filter(FinalData, sex == "gal")
gay =filter(FinalData, LGBTQ == "gay")
Druguser = filter(FinalData,DrugUser =="yes")

```


# Introduction

A new  preventative vaccine (covid91) is under test in clinical trials. Given the nature of this vaccine it is suspected that the drug may exhibit different effectiveness across the variety of subpopulations tracked. Consider population subgroups by gender( guys and gals), drug use, and sexual orientation. Write a complete report that separately considers this drug for each subgroup and your recommendations as to whether people in the subgroup should take the vaccine. Structure your report in four sections each devoted to one of the four subgroups, use null and alternative hypothesis statements and correct statistical verbage when referencing p-values in conclusions.

### Coding Details

The variables from FinalData needs to be changed from the character variables into factor variables. Subgroups had to be separated into categories, Guys, Gals, Gays, and Drugusers to see which groups would take the COVID-19 vaccine.

# Guys

### barplot

```{r}
dd2<- guy %>% group_by(infected,treatment) %>% summarize(count=n())%>% 
mutate(prcnt=count/sum(count))

# the group_by followed by summarize(count=n())

basicC <- ggplot(dd2,aes(x=infected,y=count,fill=treatment))
basicC + geom_bar(stat="identity", position = "dodge")

# Now for percentage  plot

basicC <- ggplot(dd2,aes(x=infected,y=prcnt*100,fill=treatment))
basicC + geom_bar(stat = "identity",position = "dodge")
```


### Numerical results

```{r}
table1 <- xtabs(~infected + treatment, data=guy)
table2 <- xtabs(~infected + treatment, data = guy)
rowPerc(table1)
colPerc(table1)
```


### Inferential results

```{r}
chisq.test(table1)
chisqtestGC(table1)
guysfish <- fisher.test(table1)
guysfish
```

```{r}
effect <- guysfish$p.value/(guysfish$p.value +1)
effectivenessguy = guysfish$estimate/(guysfish$estimate+1)
print(effectivenessguy*100)
```

```{r}
low=(guysfish$conf.int[1]/(guysfish$conf.int[1] +1))*100
hi=(guysfish$conf.int[2]/(guysfish$conf.int[2] +1))*100
options
(digits = 2)
```

```{r}
guysfish <- fisher.test(table1)
guysfish
```

## Effectiveness Confidence Interval `r low` to `r hi`


# Gal

### barplot

```{r}
dd2<- gal %>% group_by(infected,treatment) %>% summarize(count=n())%>% 
mutate(prcnt=count/sum(count))

# the group_by followed by summerize(count=n())

basicC <- ggplot(dd2,aes(x=infected,y=count,fill=treatment))
basicC + geom_bar(stat="identity", position = "dodge")

# Now for percentage  plot

basicC <- ggplot(dd2,aes(x=infected,y=prcnt*100,fill=treatment))
basicC + geom_bar(stat = "identity",position = "dodge")


```

### Numerical Results

```{r}
table1 <- xtabs(~infected + treatment, data=gal)
table2 <- xtabs(~infected + treatment, data=gal)
rowPerc(table1)
colPerc(table1)
```


## Inferential  

```{r}
chisq.test(table1)
chisqtestGC(table1)
galsfish <- fisher.test(table1)
galsfish
```


```{r}
effect <- galsfish$p.value/(galsfish$p.value +1)
effectivenessgal = galsfish$estimate/(galsfish$estimate+1)
print(effectivenessgal*100)
```

```{r}
low=(galsfish$conf.int[1]/(galsfish$conf.int[1] +1))*100
hi=(galsfish$conf.int[2]/(galsfish$conf.int[2] +1))*100
options
(digits = 2)
```

```{r}
galsfish <- fisher.test(table1)
galsfish
```

## Effectiveness Confidence Interval `r low` to `r hi`



# Gay Sexual Oreintation

### barplot

```{r}
dd2<- gay %>% group_by(infected,treatment) %>% summarize(count=n())%>% 
mutate(prcnt=count/sum(count))

# the group_by followed by summarize(count=n())

basicC <- ggplot(dd2,aes(x=infected,y=count,fill=treatment))
basicC + geom_bar(stat="identity", position = "dodge")

# Now for percentage  plot

basicC <- ggplot(dd2,aes(x=infected,y=prcnt*100,fill=treatment))
basicC + geom_bar(stat = "identity",position = "dodge")

```


### Numerical Results

```{r}
table1 <- xtabs(~infected + treatment, data=gay)
table2 <- xtabs(~infected + treatment, data=gay)
rowPerc(table1)
colPerc(table1)
```


### Inferential Results

```{r}
chisq.test(table1)
chisqtestGC(table1)
gaysfish <- fisher.test(table1)
gaysfish
```


```{r}
effect <- gaysfish$p.value/(gaysfish$p.value +1)
effectivenessgay = gaysfish$estimate/(gaysfish$estimate+1)
print(effectivenessgay*100)
```

```{r}
low=(gaysfish$conf.int[1]/(gaysfish$conf.int[1] +1))*100
hi=(gaysfish$conf.int[2]/(gaysfish$conf.int[2] +1))*100
options
(digits = 2)
```

```{r}
gaysfish <- fisher.test(table1)
gaysfish
```

## Effectiveness Confidence Interval `r low` to `r hi`


# Drug Users

### barplot

```{r}
dd2<- Druguser %>% group_by(infected,treatment) %>% summarize(count=n())%>% 
mutate(prcnt=count/sum(count))

# the group_by followed by summerize(count=n())

basicC <- ggplot(dd2,aes(x=infected,y=count,fill=treatment))
basicC + geom_bar(stat="identity", position = "dodge")

# Now for percentage  plot

basicC <- ggplot(dd2,aes(x=infected,y=prcnt*100,fill=treatment))
basicC + geom_bar(stat = "identity",position = "dodge")


```


### Numerical Results

```{r}
table1 <- xtabs(~infected + treatment, data=Druguser)
table2 <- xtabs(~infected + treatment, data=Druguser)
rowPerc(table1)
colPerc(table1)
```

### Inferential Results

```{r}
chisq.test(table1)
chisqtestGC(table1)
Drugusersfish <- fisher.test(table1)
Drugusersfish
```



```{r}
effect <- Drugusersfish$p.value/(Drugusersfish$p.value +1)
effectivenessdrug = Drugusersfish$estimate/(Drugusersfish$estimate+1)
print(effectivenessdrug*100)
```

```{r}
low=(Drugusersfish$conf.int[1]/(Drugusersfish$conf.int[1] +1))*100
hi=(Drugusersfish$conf.int[2]/(Drugusersfish$conf.int[2] +1))*100
options
(digits = 2)
```

```{r}
Drugusersfish <- fisher.test(table1)
Drugusersfish
```

## Effectiveness Confidence Interval `r low` to `r hi`



# Discussion and Conclusion 


